---
title: "Topical siRNA therapy of diabetic-like wound healing"
author: "Eva Neuhoferova & Marek Kindermann et al"
date: '2024-06-28'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{=html}
<style>
body {
text-align: justify}
</style>
```

### Introduction

\large

This report shows source data, statistical code and outputs of
*regression models* in a *Bayesian framework*.

Specifically, analysis for Figure 5A, B, C, D and Figure 3B is shown.

All analyses were done using [R sofware](https://www.r-project.org/) in
[RStudio](https://www.rstudio.com/) interface.

<br>

### Software version

```{r}
### Clear R environment:
rm(list=ls())
R.version
rstudioapi::versionInfo()$version
```

### Libraries

```{r warning=FALSE, message=FALSE, errors=FALSE}
### Uploading R packages:
library(here)            # file referencing
library(dplyr)           # for data manipulation
library(ggplot2)         # plot data
library(brms)            # Bayesian model fit
library(bayesplot)       # plot posterior distributions
library(loo)             # leave-one-out cross-validation
library(tidyr)           # data manipulation
library(tibble)          # related to tidyr
library(ggdist)          # plot posterior: stat_halfeye(), stat_slab()
library(wesanderson)     # set posterior palette: wes_palette()
library(marginaleffects) # get the final effect size
```

```{r}
### Uploading R packages:
citation("here")            # file referencing
citation("dplyr")           # for data manipulation
citation("ggplot2")         # plot data
citation("brms")            # Bayesian model fit
citation("bayesplot")       # plot posterior distributions
citation("loo")             # leave-one-out cross-validation
citation("tidyr")           # data manipulation
citation("tibble")          # related to tidyr
citation("ggdist")          # plot posterior: stat_halfeye(), stat_slab()
citation("wesanderson")     # set posterior palette: wes_palette()
citation("marginaleffects") # get the final effect size
```

### WOUND HEALING - Figure 5A, B and D

The regression model is applied to the time-to-event data, which are
shown in *Figure 5A, B* in the manuscript. The resulting effect size is
shown in *Figure 5D* utilizing full posterior distributions.

The *Weibull model* allows us to assume a monotonically increasing
hazard function *h_W (t)*. In other words, the chance of wound closure
increases with time. Results of the Weibull regression model in *Figure
5A-B* (smooth curves) are shown in terms of the survival function
*S_W(t)*, which describes the probability that wound closure has not
occurred by the elapsed time.

For the definitions of *h_W (t)* and *S_W (t)* see the manuscript. For a
complete discussion of statistical analysis, see the *Supporting
Information*, section *Statistical analysis*.

I found the text from Solomon Kurz as extremely helpful - see chapter
11.5 in "Statistical rethinking with brms, ggplot2, and the tidyverse:
Second edition version 0.4.0; A Solomon Kurz; 2023-01-26; Bonus:
Survival analysis". (<https://bookdown.org/content/4857/>)

### Load data:

File *Figure_5.txt* from the GitHub repository contains the source data
(see link <https://github.com/KindermannMarek/Topical-siRNA-therapy-1>).

*treat*, *group_id*, *status* should be coded as **factors**\
*day_of_scar_formation*, *censoring_status* are **numerical variables**

```{r message=FALSE}
### Load data
rm(list=ls())
### Upload (!Use your own path to the data table! Load file *Figure_5.txt* from the repository.)
file.exists("c:/Users/KindermannM/Documents/OneDrive_hotmail_gmail/4_biomedical_center/R2/1_uochb/Projects/HPMA_patent/data/Figure_4.txt")
### Upload (!Use your own path to the data table! Load file *Figure_5.txt* from the repository.)
data <- read.delim("c:/Users/KindermannM/Documents/OneDrive_hotmail_gmail/4_biomedical_center/R2/1_uochb/Projects/HPMA_patent/data/Figure_4.txt")
attach(data)

treat <- as.factor(treat)
group_ID <- as.factor(group_ID)
status <- as.factor(status)
day_of_scar_formation <- as.numeric(day_of_scar_formation)
censoring_status <- as.numeric(censoring_status)

data
```

### BM1

The weakly informative priors for the coefficients *𝛽* are centered
around 0, reflecting the conservative assumption that the difference
between the reference treatment and the others can be positive, negative
or equal to zero. Standard deviation (sd) of this prior was estimated as
sd = 2 × standard deviation(day_of_scar_formation); prior(normal(0, 6)).

See section *Statistical analysis* in manuscript for the formal model
notation.

```{r message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

data$treat <- as.factor(data$treat)
data$treat  <- relevel(data$treat, ref = "A")

bform_1 <- bf(day_of_scar_formation ~ 1 + treat)
prior_1 <- c(prior(normal(0, 6), class = b))

fit_1 <- brm(bform_1,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1 <- add_criterion(fit_1, "loo")
```

### BM1.3

Try to include the grouping variable *group_ID* to improve the model.

```{r}
### Bayesian model
set.seed = 17

data$treat <- as.factor(data$treat)
data$treat  <- relevel(data$treat, ref = "A")

bform_1.3 <- bf(day_of_scar_formation ~ 1 + treat + (1 | group_ID))
prior_1.3 <- c(prior(normal(0, 2), class = b))

fit_1.3 <- brm(bform_1.3,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1.3,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1.3 <- add_criterion(fit_1.3, "loo")
```

The model diagnostic revealed 204 divergent transitions after warmup.
Therefore, we utilize BM1 for the final analysis.

```{r}
###  Summary of a fitted model 
summary(fit_1)

### Display conditional effects
conditional_effects(fit_1) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1)

### "Leave-one-out extreme value diagnostic"
loo(fit_1, 'moment_match = TRUE')
plot(loo(fit_1, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1)

```


### BM1.1

Prior sensitivity analysis for BM1: Try to check the more narrower
prior.

```{r message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17
bform_1.1 <- bf(day_of_scar_formation ~ 1 + treat)
prior_1.1 <- c(prior(normal(0, 1), class = b))

fit_1.1 <- brm(bform_1.1,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1.1,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1.1 <- add_criterion(fit_1.1, "loo")
```

```{r}
###  Summary of a fitted model 
summary(fit_1.1)

### Display conditional effects
conditional_effects(fit_1.1) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1.1, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1.1)

### "Leave-one-out extreme value diagnostic"
loo(fit_1.1, 'moment_match = TRUE')
plot(loo(fit_1.1, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1.1) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1.1)
```

### BM1.2

Prior sensitivity analysis for BM1: Try to check the more wider prior.

```{r message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17
bform_1.2 <- bf(day_of_scar_formation ~ 1 + treat)
prior_1.2 <- c(prior(normal(0, 12), class = b))

fit_1.2 <- brm(bform_1.2,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1.2,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1.2 <- add_criterion(fit_1.2, "loo")
```

Both sensitivity analysis seem to be fine (there is no substantial
change in the estimated parameters). The results are consistent with the
proposed priors.

```{r }
###  Summary of a fitted model 
summary(fit_1.2)

### Display conditional effects
conditional_effects(fit_1.2) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1.2, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1.2)

### "Leave-one-out extreme value diagnostic"
loo(fit_1.2, 'moment_match = TRUE')
plot(loo(fit_1.2, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1.2) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1.2)
```

### BM1 - Get survival and hazard function

```{r message=FALSE, warning=FALSE}
### See chapter 11.5 in Statistical rethinking with brms, ggplot2, and the 
### tidyverse: Second edition version 0.4.0 
### A Solomon Kurz; 2023-01-26
### Bonus: Survival analysis

### Back-transformation of estimated β parameters into the λ metric 
### See equation (4) in manuscript
est_shape <- 37.85
est_scaleA <- exp(fixef(fit_1)[1, 1]) / gamma(1 + 1/est_shape)
est_scaleB <- exp(fixef(fit_1)[2, 1]) / gamma(1 + 1/est_shape)
est_scaleC <- exp(fixef(fit_1)[3, 1]) / gamma(1 + 1/est_shape)
est_scaleD <- exp(fixef(fit_1)[4, 1]) / gamma(1 + 1/est_shape)
est_scaleE <- exp(fixef(fit_1)[5, 1]) / gamma(1 + 1/est_shape)

ll_scaleA <- exp(fixef(fit_1)[1, 3]) / gamma(1 + 1/est_shape)
ll_scaleB <- exp(fixef(fit_1)[2, 3]) / gamma(1 + 1/est_shape)
ll_scaleC <- exp(fixef(fit_1)[3, 3]) / gamma(1 + 1/est_shape)
ll_scaleD <- exp(fixef(fit_1)[4, 3]) / gamma(1 + 1/est_shape)
ll_scaleE <- exp(fixef(fit_1)[5, 3]) / gamma(1 + 1/est_shape)

hh_scaleA <- exp(fixef(fit_1)[1, 4]) / gamma(1 + 1/est_shape)
hh_scaleB <- exp(fixef(fit_1)[2, 4]) / gamma(1 + 1/est_shape)
hh_scaleC <- exp(fixef(fit_1)[3, 4]) / gamma(1 + 1/est_shape)
hh_scaleD <- exp(fixef(fit_1)[4, 4]) / gamma(1 + 1/est_shape)
hh_scaleE <- exp(fixef(fit_1)[5, 4]) / gamma(1 + 1/est_shape)

### Get survival functions
### See equation (2) in manuscript
est <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(est_SA = exp(-(days/est_scaleA)^est_shape)) %>%
  mutate(est_SB = exp(-(days/est_scaleB)^est_shape)) %>%
  mutate(est_SC = exp(-(days/est_scaleC)^est_shape)) %>%
  mutate(est_SD = exp(-(days/est_scaleD)^est_shape)) %>%
  mutate(est_SE = exp(-(days/est_scaleE)^est_shape))

ll <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(ll_SA = exp(-(days/ll_scaleA)^est_shape)) %>%
  mutate(ll_SB = exp(-(days/ll_scaleB)^est_shape)) %>%
  mutate(ll_SC = exp(-(days/ll_scaleC)^est_shape)) %>%
  mutate(ll_SD = exp(-(days/ll_scaleD)^est_shape)) %>%
  mutate(ll_SE = exp(-(days/ll_scaleE)^est_shape))

hh <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(hh_SA = exp(-(days/hh_scaleA)^est_shape)) %>%
  mutate(hh_SB = exp(-(days/hh_scaleB)^est_shape)) %>%
  mutate(hh_SC = exp(-(days/hh_scaleC)^est_shape)) %>%
  mutate(hh_SD = exp(-(days/hh_scaleD)^est_shape)) %>%
  mutate(hh_SE = exp(-(days/hh_scaleE)^est_shape))

### Get hazard functions
### See equation (3) in manuscript
est_hazard <-
  data.frame(days = seq(0, 20, 0.001)) %>%
  mutate(est_SA = (est_shape/est_scaleA) * (days/est_scaleA)^(est_shape - 1)) %>%
  mutate(est_SA = est_SA / max(est_SA)) %>%
  mutate(est_SB = (est_shape/est_scaleB) * (days/est_scaleB)^(est_shape - 1)) %>%
  mutate(est_SB = est_SB / max(est_SB)) %>%
  mutate(est_SC = (est_shape/est_scaleC) * (days/est_scaleC)^(est_shape - 1)) %>%
  mutate(est_SC = est_SC / max(est_SC)) %>%
  mutate(est_SD = (est_shape/est_scaleD) * (days/est_scaleD)^(est_shape - 1)) %>%
  mutate(est_SD = est_SD / max(est_SD)) %>%
  mutate(est_SE = (est_shape/est_scaleE) * (days/est_scaleE)^(est_shape - 1)) %>% 
  mutate(est_SE = est_SE / max(est_SE)) 
```

### BM1 - Plot Bayesian posteriors showing the effect size (Figure 5D):

```{r message=FALSE, warning=FALSE}
### Plot the full posterior distributions
post <-
  as_draws_df(fit_1)  %>% 
  pivot_longer(starts_with("b_")) %>%
  group_by(.draw)  %>%
  mutate(value2 = value + value[name == "b_Intercept"]) %>%    
    mutate(value2 = case_when(
           value2 == value2[name == "b_Intercept"] ~ value[name == "b_Intercept"], 
           TRUE ~ value2  # Keep other values unchanged
  )) %>%
  mutate(days = qweibull(p = .5, shape, scale = exp(value2) / gamma( 1 + 1/shape))) %>%
  mutate(days_diff = days - days[name == "b_Intercept" ])

(g1 <-
  post %>%
  ggplot(aes(x = days, y = name)) +
  stat_halfeye(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(0.67, .89, 1),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  scale_colour_manual(values = c("#000000")) +   # posterior curve color (must be here)
  labs(x="Time until 50% of mice experienced wound closure") +
  labs(y="") +
  scale_x_continuous(limits = c(11, 22),
                     breaks = (c(10, 12, 14, 16, 18, 20, 22))) +
  theme(axis.ticks.y = element_blank()) +
  theme(legend.position="none", plot.title = element_text(size=10),
  aspect.ratio = 0.8,
  plot.margin = margin(1, 1, 1, 1, "cm"),
  axis.text=element_text(size=10),
  axis.title=element_text(size=10),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.border = element_rect(colour = "black", fill=NA, size=0.5),
  panel.grid.major = element_line(color = "grey", size = 0.5),
  panel.background = element_rect(fill = "white", colour = "white"),
  legend.background = element_rect(fill='transparent'))
)

(g2 <-
  post %>%
    filter(name != "b_Intercept") %>%
  ggplot(aes(x = days_diff, y = name)) +
  stat_halfeye(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(.89),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  scale_colour_manual(values = c("#000000")) +   # posterior curve color (must be here)
  labs(x="Time until 50% of mice experienced wound closure") +
  labs(y="") +
  scale_x_continuous(limits = c(-9, 3),
                     breaks = (c(-8, -6, -4, -2, 0, 2)))  +
  theme(axis.ticks.y = element_blank()) +
  theme(legend.position="none", plot.title = element_text(size=10),
  aspect.ratio = 0.8,
  plot.margin = margin(1, 1, 1, 1, "cm"),
  axis.text=element_text(size=10),
  axis.title=element_text(size=10),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.border = element_rect(colour = "black", fill=NA, size=0.5),
  panel.grid.major = element_line(color = "grey", size = 0.5),
  panel.background = element_rect(fill = "white", colour = "white"),
  legend.background = element_rect(fill='transparent'))
)

(g3 <-
  post %>%
  filter(name != "b_Intercept") %>%
  ggplot(aes(x = days_diff, y = name)) +
  stat_gradientinterval(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(0.67, .89, 0.99),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3,
               scale = 0.2) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  scale_colour_manual(values = c("#000000")) +   # posterior curve color (must be here)
  labs(x="Time until 50% of mice experienced wound closure") +
  labs(y="") +
  scale_x_continuous(limits = c(-9, 3),
                     breaks = (c(-8, -6, -4, -2, 0, 2)))  +
  theme(axis.ticks.y = element_blank()) +
  theme(legend.position="none", plot.title = element_text(size=10),
  aspect.ratio = 0.8,
  plot.margin = margin(1, 1, 1, 1, "cm"),
  axis.text=element_text(size=10),
  axis.title=element_text(size=10),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.border = element_rect(colour = "black", fill=NA, size=0.5),
  panel.grid.major = element_line(color = "grey", size = 0.5),
  panel.background = element_rect(fill = "white", colour = "white"),
  legend.background = element_rect(fill='transparent'))
)

```

### GLUCOSE LEVEL - Figure 5C

Data were taken from "wound healing" dataset (file *Figure_5.txt* from
the GitHub repository (see link
<https://github.com/KindermannMarek/Topical-siRNA-therapy-1>).

Glucose data:

```{r}
#create animal ID
data <- data.frame(data, seq(1, 15, 1))
data <- rename(data, animal_ID = seq.1..15..1.)

#crete new dataframe
data2 <- data.frame(data$glyc_pre, data$glyc_post, data$group_ID, data$animal_ID, data$status)
data2 <-
  data2 %>%
  pivot_longer(starts_with("data.glyc")) %>%
  rename(group_ID = data.group_ID,
         animal_ID = data.animal_ID) %>%
  filter(data.status != "healthy")


data2_check <-
  data2 %>%
  filter(name != "data.glyc_post")

```

Blood glucose level is expected to be log-normaly distributed as well as
other biological parameters like weight, triglyceride concentration atd.
The prior scale follows the link function. The log-normal family has the
identity link in brms =\> prior on usual scale (not log-scale).

### BM2

```{r message=FALSE, warning=FALSE, cache=TRUE}
set.seed = 17

data2$name <- as.factor(data2$name)
data2$name <- relevel(data2$name, ref = "data.glyc_pre")

data2$group_ID <- as.factor(data2$group_ID)
data2$animal_ID <- as.factor(data2$animal_ID)
data2$data.status <- as.factor(data2$data.status)

data2$value  <- as.numeric(data2$value)

bform_2 <- bf(value ~ 1 + name)
prior_2 <- c(prior(normal(0, 12), class = b))

fit_2 <- brm(bform_2,
            data = data2,
            family = brmsfamily("lognormal"),
            prior = prior_2,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_2 <- add_criterion(fit_2, "loo")
            
summary(fit_2)
conditional_effects(fit_2) 
pp_check(fit_2, ndraws = 50)
bayes_R2(fit_2)
loo(fit_2, 'moment_match = TRUE')
plot(loo(fit_2, 'moment_match = TRUE'))

```

Include info about *animal_ID* (repeated measurements). It is good to
note that it does not make sense to incorporate the *group_ID* to this
model (the animals were divided in three groups based on the reached
glucose level - the factor *group_ID* was created artificially -
assignment of *group_ID*).

## BM2.1

```{r}
fit_2.1 <- update(fit_2, value ~ 1 + name + (1 | animal_ID),
                  newdata = data2)

summary(fit_2.1)
conditional_effects(fit_2.1) 
pp_check(fit_2.1, ndraws = 50)
bayes_R2(fit_2.1)
loo(fit_2.1, 'moment_match = TRUE')
plot(loo(fit_2.1, 'moment_match = TRUE'))

loo(fit_2, fit_2.1)
```

Using leave one out cross-validation, we found that grouping variable
*animal_ID* in the form of random variable did not improve the model.
Therefore, we proceeded with *BM2*.

## BM2.2

Prior sensitivity analysis for BM2: Try to check the more narrower
prior.

```{r message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

data2$name <- as.factor(data2$name)
data2$name <- relevel(data2$name, ref = "data.glyc_pre")

data2$value  <- as.numeric(data2$value)

bform_2.2 <- bf(value ~ 1 + name)
prior_2.2 <- c(prior(normal(0, 2), class = b))

fit_2.2 <- brm(bform_2.2,
            data = data2,
            family = brmsfamily("lognormal"),
            prior = prior_2.2,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_2.2 <- add_criterion(fit_2.2, "loo")
            
summary(fit_2.2)
conditional_effects(fit_2.2) 
pp_check(fit_2.2, ndraws = 50)
bayes_R2(fit_2.2)
loo(fit_2.2, 'moment_match = TRUE')
plot(loo(fit_2.2, 'moment_match = TRUE'))

```

## BM2.3

Prior sensitivity analysis: Try to check the more wider prior.

```{r message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

data2$name <- as.factor(data2$name)
data2$name <- relevel(data2$name, ref = "data.glyc_pre")

data2$value  <- as.numeric(data2$value)

bform_2.3 <- bf(value ~ 1 + name)
prior_2.3 <- c(prior(normal(0, 20), class = b))

fit_2.3 <- brm(bform_2.3,
            data = data2,
            family = brmsfamily("lognormal"),
            prior = prior_2.3,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_2.3 <- add_criterion(fit_2.3, "loo")
            
summary(fit_2.3)
conditional_effects(fit_2.3) 
pp_check(fit_2.3, ndraws = 50)
bayes_R2(fit_2.3)
loo(fit_2.3, 'moment_match = TRUE')
plot(loo(fit_2.3, 'moment_match = TRUE'))

```

Both sensitivity analysis seem to be fine (there is no substantial
change in the estimated parameters). The results are consistent with the
proposed priors.

BM2 - plot the effect size (Figure 5C):

```{r warning=FALSE, message = FALSE, echo = FALSE}
### data table
pred2 <- predictions(fit_2,
                    type = "response",
                    re_formula = NA,
                    conf_level = 0.89) %>%
  posterior_draws() %>%
  group_by(drawid) %>%
  distinct(draw, .keep_all = TRUE) %>%  ### keep only unique draws and keep all variables untouched  
  mutate(diff = draw[name == "data.glyc_post"] - draw[name == "data.glyc_pre"])  %>%
  distinct(diff, .keep_all = TRUE)

p <-
pred2  %>%
      ggplot(aes(x = diff,
                 alpha = 0.8)) +
        stat_halfeye(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(.89),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "red") +
  labs(x = "estimated ddCt", y = "")
p

```

BM2 - plot the final effect size:

```{r warning=FALSE, message = FALSE, echo = FALSE}
### data table
ate2  <- comparisons(
 fit_2,
 variables = list(name = "reference"),
 by = c("name"),
 re_formula = NA, ### fixed effects only
 type = "response", ### posterior_epred methods
) %>%
  posteriordraws()


data_filt <-
  data %>%
  filter(status != "healthy")

p <-
      ggplot() +
        stat_halfeye(data = ate2,
                    aes(x = draw,
                         y = contrast,
                         alpha = 0.8,
                         fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(.89),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  geom_point(data = data_filt, 
             mapping = aes(x = diff, y = group_ID),
             position = "jitter",
             shape = 21) + 
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "red") +
  labs(x = "estimated ddCt", y = "")
p

```

### qPCR (in vitro) - Figure 3B

### Load data:

File *Figure_4B.txt* from the GitHub repository contains the source data
(see link <https://github.com/KindermannMarek/Topical-siRNA-therapy-1>).

```{r}
library(readxl)
### Upload (!Use your own path to the data table! Load file *Figure_4B.txt* from the repository.)
qpcrinvitro <- read_excel("c:\\Users\\KindermannM\\Documents\\OneDrive_hotmail_gmail\\4_biomedical_center\\R2\\1_uochb\\Projects\\HPMA_patent\\data\\qPCR_invitro\\qpcrinvitro28102024.xlsx", 
    sheet = "Sheet1")

```

Data wrangling:

```{r}
qpcrinvitro <-
  qpcrinvitro %>%
  mutate(dct = ct_TBP - ct_MMP9)
```

The study design incorporates a nested structure with technical
replicates per experiment. This hierarchical arrangement is reflected in
the multilevel model by nesting technical replicates within experiment.
The model accounts for potential variability in gene expression both
between experiments and among technical replicates within the
experiment.

Coefficients *𝛽* are centered around 0; standard deviation (sd) of this
prior was estimated as sd = 2 × standard deviation(dCt); prior(normal(0,
11)).\

### BM3

random intercept for a technical replicate nested in a given experiment:

```{r , message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

qpcrinvitro$sample <- as.factor(qpcrinvitro$sample)
qpcrinvitro$ct_TBP <- as.numeric(qpcrinvitro$ct_TBP)
qpcrinvitro$ct_MMP9 <- as.numeric(qpcrinvitro$ct_MMP9)
qpcrinvitro$techrep <- as.factor(qpcrinvitro$techrep )
qpcrinvitro$exp <- as.factor(qpcrinvitro$exp )

qpcrinvitro$sample <- relevel(qpcrinvitro$sample, ref = "ctrl")


bform_3 <- bf(dct ~ 1 + sample + (1 | exp / techrep))
prior_3 <- c(prior(normal(0, 11), class = b, coef = "samplenf"),
             prior(normal(0, 11), class = b, coef = "samplesiR"))

fit_3 <- brm(bform_3,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3 <- add_criterion(fit_3, "loo")
            
summary(fit_3)
conditional_effects(fit_3) 
pp_check(fit_3, ndraws = 50)
bayes_R2(fit_3)
loo(fit_3, 'moment_match = TRUE')
plot(loo(fit_3, 'moment_match = TRUE'))


library(emmeans)
fixef(fit_3) ### population-level (fixed) effects
ranef(fit_3) ### group-level (random) effects
coef(fit_3)  ### extract combined coefficients


```

### BM3.1

intercept (random intercept for a technical replicate nested in a given
experiment) and slope (random slope for the experiment) are calculated
independently - without any assumed correlation between them (not
calculated in the same grouping):

```{r , message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

qpcrinvitro$sample <- as.factor(qpcrinvitro$sample)
qpcrinvitro$ct_TBP <- as.numeric(qpcrinvitro$ct_TBP)
qpcrinvitro$ct_MMP9 <- as.numeric(qpcrinvitro$ct_MMP9)
qpcrinvitro$techrep <- as.factor(qpcrinvitro$techrep )
qpcrinvitro$exp <- as.factor(qpcrinvitro$exp )

qpcrinvitro$sample <- relevel(qpcrinvitro$sample, ref = "ctrl")


bform_3.1 <- bf(dct ~ 1 + sample + (1 | exp / techrep) + (0 + sample | exp))
prior_3.1 <- c(prior(normal(0, 11), class = b, coef = "samplenf"),
             prior(normal(0, 11), class = b, coef = "samplesiR"))

fit_3.1 <- brm(bform_3.1,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.1,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.1 <- add_criterion(fit_3.1, "loo")
            
summary(fit_3.1)
conditional_effects(fit_3.1) 
pp_check(fit_3.1, ndraws = 50)
bayes_R2(fit_3.1)
loo(fit_3.1, 'moment_match = TRUE')
plot(loo(fit_3.1, 'moment_match = TRUE'))

```

### BM3.2

random intercept and random slope are calculated in the same grouping:

```{r, message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

qpcrinvitro$sample <- as.factor(qpcrinvitro$sample)
qpcrinvitro$ct_TBP <- as.numeric(qpcrinvitro$ct_TBP)
qpcrinvitro$ct_MMP9 <- as.numeric(qpcrinvitro$ct_MMP9)
qpcrinvitro$techrep <- as.factor(qpcrinvitro$techrep )
qpcrinvitro$exp <- as.factor(qpcrinvitro$exp )

qpcrinvitro$sample <- relevel(qpcrinvitro$sample, ref = "ctrl")


bform_3.2 <- bf(dct ~ 1 + sample + (1 + sample | exp / techrep))
prior_3.2 <- c(prior(normal(0, 11), class = b, coef = "samplenf"),
             prior(normal(0, 11), class = b, coef = "samplesiR"))

fit_3.2 <- brm(bform_3.2,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.2,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.2 <- add_criterion(fit_3.2, "loo")
            
summary(fit_3.2)
conditional_effects(fit_3.2) 
pp_check(fit_3.2, ndraws = 50)
bayes_R2(fit_3.2)
loo(fit_3.2, 'moment_match = TRUE')
plot(loo(fit_3.2, 'moment_match = TRUE'))

```

### BM3.5

random intercept for the experiment only:

```{r , message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

qpcrinvitro$sample <- as.factor(qpcrinvitro$sample)
qpcrinvitro$ct_TBP <- as.numeric(qpcrinvitro$ct_TBP)
qpcrinvitro$ct_MMP9 <- as.numeric(qpcrinvitro$ct_MMP9)
qpcrinvitro$techrep <- as.factor(qpcrinvitro$techrep )
qpcrinvitro$exp <- as.factor(qpcrinvitro$exp )

qpcrinvitro$sample <- relevel(qpcrinvitro$sample, ref = "ctrl")


bform_3.5 <- bf(dct ~ 1 + sample + (1 | exp))
prior_3.5 <- c(prior(normal(0, 11), class = b, coef = "samplenf"),
             prior(normal(0, 11), class = b, coef = "samplesiR"))

fit_3.5 <- brm(bform_3.5,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.5,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.5 <- add_criterion(fit_3.5, "loo")
            
summary(fit_3.5)
conditional_effects(fit_3.5) 
pp_check(fit_3.5, ndraws = 50)
bayes_R2(fit_3.5)
loo(fit_3.5, 'moment_match = TRUE')
plot(loo(fit_3.5, 'moment_match = TRUE'))

```

### BM3.6

model with the fixed factor only:

```{r, message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17

qpcrinvitro$sample <- as.factor(qpcrinvitro$sample)
qpcrinvitro$ct_TBP <- as.numeric(qpcrinvitro$ct_TBP)
qpcrinvitro$ct_MMP9 <- as.numeric(qpcrinvitro$ct_MMP9)
qpcrinvitro$techrep <- as.factor(qpcrinvitro$techrep )
qpcrinvitro$exp <- as.factor(qpcrinvitro$exp )

qpcrinvitro$sample <- relevel(qpcrinvitro$sample, ref = "ctrl")


bform_3.6 <- bf(dct ~ 1 + sample )
prior_3.6 <- c(prior(normal(0, 11), class = b, coef = "samplenf"),
             prior(normal(0, 11), class = b, coef = "samplesiR"))

fit_3.6 <- brm(bform_3.6,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.6,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.6 <- add_criterion(fit_3.6, "loo")
            
summary(fit_3.6)
conditional_effects(fit_3.6) 
pp_check(fit_3.6, ndraws = 50)
bayes_R2(fit_3.6)
loo(fit_3.6, 'moment_match = TRUE')
plot(loo(fit_3.6, 'moment_match = TRUE'))

```

Comparison of models. The results of the cross-validation show that
models BM3, BM3.1 and BM3.2 are comparable and outperform models BM3.5
and BM3.6.

Modeling the random slope for models BM3.1 and BM3.2 represents
additional complexity compared to model BM3 (BM3 includes only random
intercept) which does not bring noticeable improvement. Thus, let's
stick to our initial expectation (model BM3 model).

```{r}
loo(fit_3, fit_3.1, fit_3.2, fit_3.5, fit_3.6)
```

Let's take a look at the group-level estimates (just out of interest):

```{r}
pred <- predictions(fit_3,
    newdata = datagrid(sample = c("ctrl", "nf", "siR"), 
                       exp = unique),
    re_formula = ~(1|exp / techrep),
    allow_new_levels = TRUE,
    type = "response", ### posterior_epred method
    by = c("sample", "exp")) 

pred <-  posterior_draws(pred)
pred %>%
        ggplot(aes(y = draw,
                   x = sample,
                   alpha = 1)) +
        # stat_halfeye() +
  stat_pointinterval(.width = c(0.95),
                    point_interval = "median_qi")+
  theme_bw() +
  theme(legend.position = "none") +
  facet_grid(~exp) + 
  labs(x = "dCt", y = "treatment")

```

Plot the final effect size (population-level estimates). Let's try
different implementations first to see that everything is ok:

```{r}

### [1]
conditional_effects(fit_3, conditions = NULL, re_formula = NA, method = "posterior_epred", prob = 0.89, ask = FALSE)

### [2]
plot_predictions(fit_3, condition = "sample", type = "response", re_formula = NA, conf_level = 0.89)

### [3]
pred1 <- predictions(fit_3,
                    type = "response",
                    re_formula = NA,
                    conf_level = 0.89)
pred1 <- posterior_draws(pred1)
ggplot(pred1, aes(y = draw, x = sample)) +
    stat_pointinterval(.width = c(0.89))

```

Plot the full posterior distributions (Figure 4B):

```{r}
pred1 <- predictions(fit_3,
                     type = "response",
                     re_formula = NA,
                     conf_level = 0.89)
pred1 <- posterior_draws(pred1)
ggplot(pred1, aes(y = draw, x = sample)) +
stat_pointinterval(.width = c(0.89))

pred1_compare <-
  pred1 %>%
  group_by(drawid, exp, sample) %>%
  distinct(draw, .keep_all = TRUE) %>%
  ungroup() %>%
  group_by(drawid, exp) %>%
  mutate(diff = draw - draw[sample == "ctrl"]) %>%
  filter(sample != "ctrl") %>%
  ungroup()


pred1_compare %>%
      ggplot(aes(x = diff,
                 y = sample,
                 alpha = 0.8)) +
        stat_halfeye(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(.89),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "red") +
  labs(x = "estimated ddCt", y = "")



```

### BM3.7

Prior sensitivity analysis for BM3: Try to check the more narrower
prior.

```{r , message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17


bform_3.7 <- bf(dct ~ 1 + sample + (1 | exp / techrep))
prior_3.7 <- c(prior(normal(0, 2), class = b, coef = "samplenf"),
             prior(normal(0, 2), class = b, coef = "samplesiR"))

fit_3.7 <- brm(bform_3.7,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.7,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.7 <- add_criterion(fit_3.7, "loo")
            
summary(fit_3.7)
conditional_effects(fit_3.7) 
pp_check(fit_3.7, ndraws = 50)
bayes_R2(fit_3.7)
loo(fit_3.7, 'moment_match = TRUE')
plot(loo(fit_3.7, 'moment_match = TRUE'))


library(emmeans)
fixef(fit_3.7) ### population-level (fixed) effects
ranef(fit_3.7) ### group-level (random) effects
coef(fit_3.7)  ### extract combined coefficients


```

### BM3.8

Prior sensitivity analysis for BM3: Try to check the more wider prior.

```{r , message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17


bform_3.8 <- bf(dct ~ 1 + sample + (1 | exp / techrep))
prior_3.8 <- c(prior(normal(0, 20), class = b, coef = "samplenf"),
             prior(normal(0, 20), class = b, coef = "samplesiR"))

fit_3.8 <- brm(bform_3.8,
            data = qpcrinvitro,
            family = brmsfamily("normal"),
            prior = prior_3.8,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_3.8 <- add_criterion(fit_3.8, "loo")
            
summary(fit_3.8)
conditional_effects(fit_3.8) 
pp_check(fit_3.8, ndraws = 50)
bayes_R2(fit_3.8)
loo(fit_3.8, 'moment_match = TRUE')
plot(loo(fit_3.8, 'moment_match = TRUE'))


library(emmeans)
fixef(fit_3.8) ### population-level (fixed) effects
ranef(fit_3.8) ### group-level (random) effects
coef(fit_3.8)  ### extract combined coefficients


```

Both sensitivity analysis seem to be fine (there is no substantial
change in the estimated parameters). The results are consistent with the
proposed priors.
