---
title: "Topical siRNA therapy of diabetic-like wound healing"
author: "Eva Neuhoferova & Marek Kindermann et al"
date: '2024-06-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
body {
text-align: justify}
</style>  

### Introduction
\large
This report shows source data, statistical code and outputs of *Weibull regression model* in a *Bayesian framework*. The regression model is applied to the time-to-event data, which are shown in *Figure 4A, B* in the manuscript. The resulting effect size is shown in *Figure 4D* utilizing full posterior distributions.  

The *Weibull model* allows us to assume a monotonically increasing hazard function *h_W (t)*. In other words, the chance of wound closure increases with time. Results of the Weibull regression model in *Figure 4A-B* (smooth curves) are shown in terms of the survival function *S_W (t)*, which describes the probability that wound closure has not occurred by the elapsed time.

For the definitions of *h_W (t)* and *S_W (t)* see the manuscript.

For a complete discussion of statistical analysis, see the *Supporting Information*, section *Statistical analysis*. 

All analyses were done using [R sofware](https://www.r-project.org/) in [RStudio](https://www.rstudio.com/) interface.

<br>

### Software version
```{r clear}
### Clear R environment:
rm(list=ls())
R.version
rstudioapi::versionInfo()$version
```

### Libraries
```{r libraries, warning=FALSE, message=FALSE, errors=FALSE}
### Uploading R packages:
library(here)           # file referencing
library(dplyr)          # for data manipulation
library(ggplot2)        # plot data
library(brms)           # Bayesian model fit
library(bayesplot)      # plot posterior distributions
library(loo)            # leave-one-out cross-validation
library(tidyr)          # data manipulation
library(tibble)         # related to tidyr
library(ggdist)         # plot posterior: stat_halfeye(), stat_slab()
library(wesanderson)    # set posterior palette: wes_palette()
```

### Load data
*treat*, *group_id*, *status* should be coded as **factors**   
*day_of_scar_formation*, *censoring_status* are **numerical variables**
```{r load data, message=FALSE}
### Load data
rm(list=ls())
### Upload (! Use your own path to the data table !)
file.exists("c:/Users/KindermannM/Documents/OneDrive_hotmail_gmail/4_biomedical_center/R2/1_uochb/Projects/HPMA_patent/data/Figure_4.txt")
### Upload (! Use your own path to the data table !)
data <- read.delim("c:/Users/KindermannM/Documents/OneDrive_hotmail_gmail/4_biomedical_center/R2/1_uochb/Projects/HPMA_patent/data/Figure_4.txt")
attach(data)

treat <- as.factor(treat)
group_ID <- as.factor(group_ID)
status <- as.factor(status)
day_of_scar_formation <- as.numeric(day_of_scar_formation)
censoring_status <- as.numeric(censoring_status)

data
```

### Bayesian model 1 (BM1)
The weakly informative priors for the coefficients *ð›½* are centered around 0, reflecting the conservative assumption that 
the difference between the reference treatment and the others can be positive, negative or equal to zero. Standard
deviation (sd) of this prior was estimated as sd = 2 Ã— standard deviation(day_of_scar_formation); prior(normal(0, 6)).  
See section *Statistical analysis* in manuscript for the formal model notation. 
```{r BM1, message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17
bform_1 <- bf(day_of_scar_formation ~ 0 + treat)
prior_1 <- c(prior(normal(0, 6), class = b))

fit_1 <- brm(bform_1,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1 <- add_criterion(fit_1, "loo")
```
```{r BM1 diagnostic}
###  Summary of a fitted model 
summary(fit_1)

### Display conditional effects
conditional_effects(fit_1) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1)

### "Leave-one-out extreme value diagnostic"
loo(fit_1, 'moment_match = TRUE')
plot(loo(fit_1, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1)

```

### Bayesian model 1.1 (BM1.1)
Prior sensitivity analysis: Try to check the more narrower prior.
```{r BM1.1, message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17
bform_1.1 <- bf(day_of_scar_formation ~ 0 + treat)
prior_1.1 <- c(prior(normal(0, 1), class = b))

fit_1.1 <- brm(bform_1.1,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1.1,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1.1 <- add_criterion(fit_1.1, "loo")
```
```{r BM1.1 diagnostic}
###  Summary of a fitted model 
summary(fit_1.1)

### Display conditional effects
conditional_effects(fit_1.1) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1.1, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1.1)

### "Leave-one-out extreme value diagnostic"
loo(fit_1.1, 'moment_match = TRUE')
plot(loo(fit_1.1, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1.1) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1.1)
```

### Bayesian model 1.2 (BM1.2)
Prior sensitivity analysis: Try to check the more wider prior.
```{r BM1.2, message=FALSE, warning=FALSE, cache=TRUE}
### Bayesian model
set.seed = 17
bform_1.2 <- bf(day_of_scar_formation ~ 0 + treat)
prior_1.2 <- c(prior(normal(0, 12), class = b))

fit_1.2 <- brm(bform_1.2,
            data = data,
            family = brmsfamily("weibull"),
            prior = prior_1.2,
            warmup = 2000,
            iter = 5000,
            chains = 4,
            cores = 4,
            control = list(adapt_delta = 0.95, max_treedepth = 15))
            fit_1.2 <- add_criterion(fit_1.2, "loo")
```
```{r BM1.2 diagnostic}
###  Summary of a fitted model 
summary(fit_1.2)

### Display conditional effects
conditional_effects(fit_1.2) 

### Distribution of outcome vs. newly simulated data distribution
### (dark line should be close to light lines)
pp_check(fit_1.2, ndraws = 50)

### Bayesian version of R-squared for regression model
bayes_R2(fit_1.2)

### "Leave-one-out extreme value diagnostic"
loo(fit_1.2, 'moment_match = TRUE')
plot(loo(fit_1.2, 'moment_match = TRUE'))

### Trace rank plot: these histograms should be reasonable uniform, with no
### chain spiking above or below the others.
mcmc_rank_overlay(fit_1.2) 

### Check whether the priors were informative: 
### Is the posterior standard deviation greater than 10% of the prior 
### standard deviation?
bayestestR::check_prior(fit_1.2)
```

### Get survival and hazard function - BM1
```{r get survival and hazard function - BM1, message=FALSE, warning=FALSE}
### See chapter 11.5 in Statistical rethinking with brms, ggplot2, and the 
### tidyverse: Second edition version 0.4.0 
### A Solomon Kurz; 2023-01-26
### Bonus: Survival analysis

### Back-transformation of estimated Î² parameters into the Î» metric 
### See equation (4) in manuscript
est_shape <- 37.85
est_scaleA <- exp(fixef(fit_1)[1, 1]) / gamma(1 + 1/est_shape)
est_scaleB <- exp(fixef(fit_1)[2, 1]) / gamma(1 + 1/est_shape)
est_scaleC <- exp(fixef(fit_1)[3, 1]) / gamma(1 + 1/est_shape)
est_scaleD <- exp(fixef(fit_1)[4, 1]) / gamma(1 + 1/est_shape)
est_scaleE <- exp(fixef(fit_1)[5, 1]) / gamma(1 + 1/est_shape)

ll_scaleA <- exp(fixef(fit_1)[1, 3]) / gamma(1 + 1/est_shape)
ll_scaleB <- exp(fixef(fit_1)[2, 3]) / gamma(1 + 1/est_shape)
ll_scaleC <- exp(fixef(fit_1)[3, 3]) / gamma(1 + 1/est_shape)
ll_scaleD <- exp(fixef(fit_1)[4, 3]) / gamma(1 + 1/est_shape)
ll_scaleE <- exp(fixef(fit_1)[5, 3]) / gamma(1 + 1/est_shape)

hh_scaleA <- exp(fixef(fit_1)[1, 4]) / gamma(1 + 1/est_shape)
hh_scaleB <- exp(fixef(fit_1)[2, 4]) / gamma(1 + 1/est_shape)
hh_scaleC <- exp(fixef(fit_1)[3, 4]) / gamma(1 + 1/est_shape)
hh_scaleD <- exp(fixef(fit_1)[4, 4]) / gamma(1 + 1/est_shape)
hh_scaleE <- exp(fixef(fit_1)[5, 4]) / gamma(1 + 1/est_shape)

### Get survival functions
### See equation (2) in manuscript
est <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(est_SA = exp(-(days/est_scaleA)^est_shape)) %>%
  mutate(est_SB = exp(-(days/est_scaleB)^est_shape)) %>%
  mutate(est_SC = exp(-(days/est_scaleC)^est_shape)) %>%
  mutate(est_SD = exp(-(days/est_scaleD)^est_shape)) %>%
  mutate(est_SE = exp(-(days/est_scaleE)^est_shape))

ll <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(ll_SA = exp(-(days/ll_scaleA)^est_shape)) %>%
  mutate(ll_SB = exp(-(days/ll_scaleB)^est_shape)) %>%
  mutate(ll_SC = exp(-(days/ll_scaleC)^est_shape)) %>%
  mutate(ll_SD = exp(-(days/ll_scaleD)^est_shape)) %>%
  mutate(ll_SE = exp(-(days/ll_scaleE)^est_shape))

hh <-
  data.frame(days = seq(0, 30, 0.001)) %>%
  mutate(hh_SA = exp(-(days/hh_scaleA)^est_shape)) %>%
  mutate(hh_SB = exp(-(days/hh_scaleB)^est_shape)) %>%
  mutate(hh_SC = exp(-(days/hh_scaleC)^est_shape)) %>%
  mutate(hh_SD = exp(-(days/hh_scaleD)^est_shape)) %>%
  mutate(hh_SE = exp(-(days/hh_scaleE)^est_shape))

### Get hazard functions
### See equation (3) in manuscript
est_hazard <-
  data.frame(days = seq(0, 20, 0.001)) %>%
  mutate(est_SA = (est_shape/est_scaleA) * (days/est_scaleA)^(est_shape - 1)) %>%
  mutate(est_SA = est_SA / max(est_SA)) %>%
  mutate(est_SB = (est_shape/est_scaleB) * (days/est_scaleB)^(est_shape - 1)) %>%
  mutate(est_SB = est_SB / max(est_SB)) %>%
  mutate(est_SC = (est_shape/est_scaleC) * (days/est_scaleC)^(est_shape - 1)) %>%
  mutate(est_SC = est_SC / max(est_SC)) %>%
  mutate(est_SD = (est_shape/est_scaleD) * (days/est_scaleD)^(est_shape - 1)) %>%
  mutate(est_SD = est_SD / max(est_SD)) %>%
  mutate(est_SE = (est_shape/est_scaleE) * (days/est_scaleE)^(est_shape - 1)) %>% 
  mutate(est_SE = est_SE / max(est_SE)) 
```

### Plot Bayesian posteriors - BM1
```{r plot Bayesian posteriors - BM1, message=FALSE, warning=FALSE}
### Plot the full posterior distributions
post <-
  as_draws_df(fit_1)  %>% 
  pivot_longer(starts_with("b_")) %>%
  mutate(days = qweibull(p = .5, shape, scale = exp(value) / gamma( 1 + 1/shape))) %>%
  mutate(days_diff = days - days[name == "b_treatC" ])

(g1 <-
  post %>%
  ggplot(aes(x = days, y = name)) +
  stat_halfeye(aes(fill = after_stat(level)),    # area under curve colors
               point_interval = mean_qi,         # circle = mean + quantile intervals
               .width = c(0.67, .89, 1),         # areas under curve
               slab_color = "#000000",           # posterior curve color
               slab_linewidth = 0.7,
               height = 3) +
               stat_spike(at = c(mean)) +
  scale_fill_brewer(type = "seq",                # collor pallete - area under curve
                    palette = "Greys",
                    direction = 1,
                    aesthetics = "fill") +
  scale_colour_manual(values = c("#000000")) +   # posterior curve color (must be here)
  labs(x="Time until 50% of mice experienced wound closure") +
  labs(y="") +
  scale_x_continuous(limits = c(11, 22),
                     breaks = (c(10, 12, 14, 16, 18, 20, 22))) +
  theme(axis.ticks.y = element_blank()) +
  theme(legend.position="none", plot.title = element_text(size=10),
  aspect.ratio = 0.8,
  plot.margin = margin(1, 1, 1, 1, "cm"),
  axis.text=element_text(size=10),
  axis.title=element_text(size=10),
  plot.background = element_rect(fill='transparent', color=NA),
  panel.border = element_rect(colour = "black", fill=NA, size=0.5),
  panel.grid.major = element_line(color = "grey", size = 0.5),
  panel.background = element_rect(fill = "white", colour = "white"),
  legend.background = element_rect(fill='transparent'))
)
```

